---
title: "Data Analysis on Food Supply"
date: "November 13, 2024"
author: "Eduardo Ramirez"
execute:
  keep-md: true
  warning: false
format:
  html:
    code-fold: true
---

# Introduction

This analysis examines two datasets: the availability of fruits and vegetables and the average protein supply over a three-year average. Both datasets provide insights into the nutritional supply at the national level, measured in grams per person per day.

## Data Sources and Quality

### 1. Availability of Fruits and Vegetables

**Source:** FAOSTAT\
**Description:** Amounts of fruits and vegetables available in a country's food supply (grams per person per day).\
**Quality:** High-quality data sourced from FAO's Food Balance Sheets, though it does not guarantee actual consumption.

### 2. Average Protein Supply (3-Year Average)

**Source:** FAOSTAT\
**Description:** Grams of protein per person per day available in a country's food supply, calculated as a three-year average.\
**Quality:** High-quality data with consideration of national-level variations.

```{r, error=FALSE, message=FALSE, warning=FALSE}
# libraries

if (!require("pacman")) install.packages("pacman")
pacman::p_load(tidyverse, sf, leaflet, maps)

fruits_vegetables <- read_csv("availability-of-fruits-and-vegetables.csv")
protein_supply <- read_csv("average-protein-supply-3-year-average.csv")


```

```{r, error=FALSE, message=FALSE, warning=FALSE}
# fruits and vegetables data
fruits_vegetables_clean <- fruits_vegetables |>
  filter(!is.na(Value)) |>
  rename(fruits_vegetables_supply = Value) |>
  select(`Region Name`, `Start Year`, fruits_vegetables_supply)

# protein supply data
protein_supply_clean <- protein_supply |>
  filter(!is.na(Value)) |>
  rename(protein_supply = Value) |>
  select(`Region Name`, `Start Year`, protein_supply)

# merge df
merged_data <- merge(fruits_vegetables_clean, protein_supply_clean, by = c("Region Name", "Start Year"))


```

# Interactive Map

```{r, error=FALSE, warning=FALSE, message=FALSE}


# Load & prep world map data
world <- st_as_sf(maps::map("world", plot = FALSE, fill = TRUE))
merged_data$`Region Name` <- as.character(merged_data$`Region Name`)
world$ID <- as.character(world$ID)
world_data <- world |>
  left_join(merged_data, by = c("ID" = "Region Name"))


# interactive map
leaflet(world_data) |>
  addTiles() |>
  addPolygons(
    fillColor = ~colorQuantile("YlGnBu", `fruits_vegetables_supply`)(`fruits_vegetables_supply`),
    weight = 1,
    opacity = 0.7,
    color = 'white',
    dashArray = '3',
    fillOpacity = 0.7,
    highlightOptions = highlightOptions(
      weight = 3,
      color = '#666',
      dashArray = '',
      fillOpacity = 0.7,
      bringToFront = TRUE
    ),
    label = ~paste(`ID`,
      "Fruits & Vegetables: ", round(`fruits_vegetables_supply`, 2), " g/capita/day",
      "Protein: ", round(`protein_supply`, 2), " g/capita/day"
    ),
    labelOptions = labelOptions(
      style = list("font-weight" = "normal", padding = "3px 8px"),
      textsize = "15px",
      direction = "auto"
    )
  ) |>
  addLegend(
    pal = colorQuantile("YlGnBu", NULL),
    values = ~`fruits_vegetables_supply`,
    opacity = 0.7,
    title = "Fruits & Vegetables (g/capita/day)",
    position = "bottomright"
  )


```
